---
title: "Prediction Assignment Writeup"
author: "Han Zhao"
date: "5/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
The goal of your project is to predict the manner in which 6 participants did the exercise.The machine learning algorithm described here is applied to the 20 test cases in the test data.

## Load data sets
```{r}
training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
                    na.strings = c("NA","#DIV/0!",""))
testing <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
                    na.strings = c("NA","#DIV/0!",""))
```

## Clean training data set and remove invalid predictors
```{r}
library(caret)
# Remove columns with Near Zero Values
training_clean <- training[ , names(training)[!(nzv(training, saveMetrics = T)[, 4])]]
# Remove columns is empty
training_clean <- training_clean[, names(training_clean)[sapply(training_clean, function(x)!(any(is.na(x)|x=="")))]]
# Remove Col 1 (X) and Col 5 (cvtd_timestamp)
training_clean <- training_clean[, -c(1,5)]
```

## Separate Training data set for Cross Validation
```{r}
set.seed(1234)
inTrain <- createDataPartition(training_clean$classe, p = 0.7, list = F)
train <- training_clean[inTrain, ]
valid <- training_clean[-inTrain, ]
```

## Prediction Model
Using the test data set and random forest method to create a prediction model.
```{r}
controlRF <- trainControl(method="cv", number=3, verboseIter = F)
fit <- train(classe ~., method = "rf", data = train, trControl=controlRF)
```

## Assuracy of the Prediction Model
Using the train data to create a prediction.
```{r}
predTrain <- predict(fit, train)
```
Measure its accuracy
```{r}
rfMatrix <- confusionMatrix(predTrain, train$classe)
```

Plot the matrix
```{r}
plot(rfMatrix$table, col = rfMatrix$byClass, 
     main = paste("Random Forest - Accuracy =",
                  rfMatrix$overall['Accuracy']))
```

## Cross Validation
Using the validation subset and create a prediction.
```{r}
predValid <- predict(fit, valid)
```
Measure its accurancy
```{r}
rfMatrix <- confusionMatrix(predValid, valid$classe)
```
Plot the Matrix
```{r}
plot(rfMatrix$table, col = rfMatrix$byClass, 
     main = paste("Random Forest - Accuracy =",
                  rfMatrix$overall['Accuracy']))
```

From the validation subset, the accuracy is very high, at above 99%. 

## Final Model
Given the level of accuracy, we will use the model above constructed by the random forest method for the prediction.
The following is the final model:
```{r}
fit$finalModel
```
The following is the list of important predictors of the model
```{r}
varImp(fit)
```

## Application of the prediction model on test data set
```{r}
predict(fit, testing)
```



